The Galactic Chemical Evolution of Carbon: Implications for Stellar Nucleosynthesis

### 2025-03-11 carbon_paper/tex/main.tex
draft for collab. comments

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Basic setup. Most papers should leave these options alone.
\documentclass[fleqn,
% referee, % this line changes to double-spaced 1 column
usenatbib]{mnras}

@@

\newcommand{\fruity}{\texttt{\hyperlink{fruity}{Fruity}}}
\newcommand{\nugrid}{\texttt{\hyperlink{nugrid}{NuGrid}}}
\newcommand{\monash}{\texttt{\hyperlink{monash}{Monash}}}
\newcommand{\aton}{\texttt{\hyperlink{aton}{ATON}}}
\newcommand{\cfactor}{2.40}
\newcommand{\nsubgiants}{14,066}

@@

Predictions of C yields from stellar evolution and supernova (SN) models are riddled with uncertainties (see, e.g., the reviews by \citealt{romano+10, KL14}).
Some salient unknowns include mass loss rates \citep{sukhbold+16, beasor+2020}, rotational mixing \citep{frischknecht+16, LC18}, nuclear reaction rates \citep{herwig+05, herwig+austin2004}, and convection \citep{chieffi2001, ventura+13}.
Recent work additionally suggests that small changes in a stars's initial mass of 
 $\sim$$0.01 M_\odot$  have significant effects on the evolution and explosion outcomes of massive stars \citep{bruenn+2023, vartanyan_burrows2023}.
 Furthermore, many stars are found in binary systems \citep{sana+12}, the effects of which are almost entirely
unexplored in stellar nucleosynthesis models \citet{farmer+21}.
Accurate predictions of metal yields from stellar evolution models remain elusive.

@@

 % Folowing many previous works  (e.g. \citealt{DTS78, BF06, prantzos+18, berg+19}; see also the recent reviews by \citealt{romano22} and \citealt{RM21}.) this paper focuses on the evolution of C in GCE models.
\citet{james+24} recently argued that the enrichment history of the Galactic disk can be described as an equilibrium phenomenon.
In equilibrium GCE, trends in abundance ratios closely trace yield ratios (see their section 5.2.2).
\citet{james+23} had previously shown that the correlation between gas-phase N and O abundances (see their Fig. 1) could be described as a consequence of the metallicity dependence of the relative N and O yields.
Under their equilibrium scenario, this argument should extend to any element whose enrichment by stellar populations does not require long timescales ($\gtrsim$few Gyr).
We show in section~\ref{sec:agb} that the characteristic delay times of C production should be $\sim$ 1 Gyr, within this limit. 
\dbnote{This is complicated by the difference between high alpha and low alpha, and 1-2Msun stars do produce carbon.... Also, I would like to call out reviews and general works on C at some point.}

@@

% \footnotetext{\strike{By metallicity, we mean the (mass) fraction of any element which is not H or He, denoted by $Z$. For the sun, we take $Z_\odot=0.0176$.} }

@@

Fig.~\ref{fig:subgiants} shows out subgiant sample plotted in the \caah\ and \caafe{} planes.\footnotemark{} In this sample, [C/Mg] increases with [Mg/H], and [C/Mg] decreases with [Mg/Fe] at fixed [Mg/H]. Our aim is to reproduce these trends.

@@

	\begin{tabular}{l l l l l}
		\hline
         & $Z_{X,\,\sun}$ & $y_X^{\rm cc}$ & $\y_X^{\rm agb}$ & $y_X^{\rm ia}$  \\
		\hline
        C & $33.9\times10^{-4}$ & Eq.~\ref{eq:y_cc} & $\cfactor\times$\fruity &  0 \\
        O & $73.3\times10^{-4}$ & $71.3\times10^{-4}$ & 0 & 0 \\
        Mg & $6.71\times10^{-4}$ & $6.52\times 10^{-4}$ & 0 & 0 \\
        Fe & $13.7\times10^{-4}$ & $4.73\times10^{-4}$ & 0 & $7.82\times10^{-4}$ \\
        N & $10.4\times10^{-4}$ & $5\times10^{-4}$ & $5\times10^{-4}M\left(\frac{Z}{Z_\odot}\right)$ & 0\\
		\hline
	\end{tabular}
\end{table}

@@

\begin{table*}
	\centering
    \caption[]{For each \agb\ yield set, the IMF-averaged \agb\ C yield at solar metallicity $y_{\rm C, 0}^{\rm AGB}$, the linear fit solar yield $\zetao$ and slope $\zetai$ (see Eq.~\ref{eq:agb_z_approx}), the fraction of solar C produced in the model $f_\odot^{\rm AGB}$, and the masses and metallicities each yield table is sampled on.
    $f_{\rm odot}^{\rm AGB}$ is calculated based on $\zetao$ assuming a total C yield of $\Yct = 0.00275$.
    }

	\label{tab:agb}
    \begin{tabular}{c  ccc  c p{4cm} p{4cm}} % four columns, alignment for each
		\hline 
        \agb\ table 
                & $y_{\rm C, \odot}^{\rm AGB}\times10^4$ %& y
                & $\zetao^{\rm AGB}\times 10^4$ %& y
                & $\zetai^{\rm AGB}\times 10^4$
                &  $f_\odot^{\rm AGB}$
                & masses ($\Mo$) & metallicities ($Z$)\\
        \hline
        \fruity 
                & 3.11
                &  $3.7\pm0.3$
                & $-3.5\pm0.3$
                & 0.13
                & 1.3, 1.5, 2, 2.5, 3, 4, 5, 6
                & 0.0001, 0.0003, 0.001, 0.002, 0.003, 0.006, 0.008, 0.01, 0.014, 0.02
                \\
        \aton 
                & -0.009
                & $1.0\pm1.8$
                & $-10. \pm 3$
                & 0.04
                & 1.5, 2, 2.5, 3, 3.5, 4, 4.5, 5, 6, 6.5, 7
                & 0.0003, 0.001, 0.002, 0.004, 0.008, 0.014, 0.04
                \\
        \monash 
                &  2.96
                & $2.3 \pm 0.5$
                & $-11.0\pm 1.0$
                & 0.08
                & 1, 1.25, 1.5, 1.75, 2.25, 2.5, 2.75, 3, 3.25, 3.5, 3.75, 4, 4.5, 5, 5.5, 6, 7 
                & 0.0028, 0.007, 0.014, 0.03
                \\
        \nugrid 
                & 9.25
                & $8.1\pm 1.9$
                & $-4.6\pm1.7$
                & 0.29
                & 1, 1.65, 2, 3, 4, 5, 6, 7
                &  0.0001, 0.001, 0.006, 0.01, 0.02
                \\
		\hline
	\end{tabular}
\end{table*}

@@

Fig.~\ref{fig:y_cc} shows the integrated yields of several massive star yield models from the literature. 
\cc{} models predict a wide range of C yields, spanning nearly a factor of ten. 
 The \cite{LC18} models, which include rotation, show that the induced mixing (e.g. \citealt{frischknecht+16}) can dramatically increase the magnitude and metallicity dependence of $\Ycc$. As we will later emperically show, \cc\ C production needs to be strongly metallicity-dependent at $Z \approx \Zo$, which is consistent with \citepos{LC18} rapidly rotating models and to a lesser extent \citet{NKT13}.
The left panel of Fig.~\ref{fig:y_cc} also shows the IMF-averaged \agb\ star yields from the \fruity\ model (in Fig.~\ref{fig:agb-ssp}). Especially at $Z\approx Z_\odot$, most \cc\ models dominate \agb\ C production.

@@

\subsection{Supernova Type Ia}
The population averaged SN Ia yield, $y_{X}^{\rm ia}$, is given by,
\begin{equation}
    y_{X}^{\rm Ia} = m_{X}^{\rm Ia} \frac{N_{\rm Ia}}{M_\star}
\end{equation}
where $m_{X}^{\rm Ia}$ is the average mass of $X$ produced in a single SN Ia event, and $N_{\rm Ia} / M_\star$ is the number of such events per unit mass of star formation. \dbstrike{The latter can be expressed as an integral over the SN Ia delay-time-distribution (DTD), which describes the rate at which stellar populations produce X through SN Ia in \VICE.} Following \citet{james+23}, we use a $t^{-1.1}$  delay time distribution (DTD) with a minimum delay-time of 150 Myr, based on comparisons of the SN Ia rate as a function of redshift and the cosmic SFR (e.g. \citealt{maoz+12}).

Supernovae type Ia produce negligible C. For our adopted yield scale, the highest magnitude CCSNe yield of the models from iwamoto99 is $1.1\times10^{-4}$, corresponding to about 4\% of our total C production, however most models predict a \ia\ C yield of less than $0.4 \times 10^{-4}$. We do check that including this contribution negligibly impacts our results, so we neglect SNIa C production for simplicity.

Novae are important contributes of \C[13], but produce far less \C[12] \dbadd{CITE}.

@@

As the strength of outflows controls the resulting equilibrium $\alpha$-element abundances, we extend \JJ and create a metallicity gradient by defining
\begin{equation}\label{eq:mass_loading}
\eta(R) = \frac{y_{\alpha}^{\rm cc}}{Z_{\alpha}(R)} -1 + r %+ \tau_{\star} / \tau_{\rm sfh} 
\end{equation}
where 
\begin{equation}
    \log Z_{\alpha}(R) = \log Z_{\alpha,\ \odot} + 
    0.29 + 
    \begin{cases}
        -0.015(R-5) & R < 5 \\
        -0.09(R-5) & R \geq 5
    \end{cases}
\end{equation}
is adapted from \citet{hayden+14}.  We approximate $r=0.4$. %, and $\tau_\star/\tau_{\rm sfh} = 0.0$ .
This choice of $\eta(R)$ results in a [$\alpha$/H] gradient consistent with Milky Way observations \citep[e.g.][]{hayden+14, weinberg+19, frinchaboy+13}.
If we change our assumed $y_{\rm Mg}$, the values of $\eta$ will change similarly to maintain the correct chemical trends.

@@

In GCE, total elemental abundances are (to first order) linear with respect to combinations of yields \citep[e.g.][]{WAF17}.
\dbnote{might be useful to include a complete set of differential equations for our model somewhere to better illustrate this.}
Assume that the yields do not depend on the yields (e.g. metallicity dependence is negligable). 
Given a set of yields $\{y_X^i\}_i$ and resulting abundance evolution $\{Z_{X}^i(R, t) \}$, let $y_X'$, the yield of X, be a linear combination of $\{y_X^i\}$
\begin{equation}
    y_X' = \sum_i \alpha_i\, y_X^i \quad {\rm where\ } \alpha_i \in \mathbb{R}.
\end{equation}
Then the abundance evolution of element $X$ is
\begin{equation}\label{eq:lin_combo}
    Z_X'(R, t) = \sum_i \alpha_i Z_X^i(R, t).
\end{equation}
Therefore, we can efficiently explore combinations of yields without the need for full simulations. This allows for direct observational inference of the yield combinations which best agree with the data.
We use Monte Carole Markov Chain (MCMC) inference to generate the posterior distributions of yield coefficients. Note that many other GCE parameters are nonlinear with abundance and require full multi-zone simulations, such as star formation history, outflows, and migration prescriptions. 
Also note that metallicity-dependent yields also introduce a second-order effect (metallicity depends on yields which depend on metallicity).

Our first step is to run a multi-zone model tracking the abundances of each process we are interested in. 
We focus on the binned \textit{mean} abundances of the low-$\alpha$ sequence in \caah, and the $-0.15 \leq {\rm [M/H]} \leq -0.05$ metallicity slice in \caafe\ (with 20 bin edges between -0.4 and 0.3 in [Mg/H] and 12 bin edges between 0.025 and 0.25 in [Mg/Fe]).
We separate the C yield  into four different components.
\begin{enumerate}
    \item $\alpha$: Prefactor for an AGB yield table
    \item $\zetao$: Constant CCSN yield  ($y_0 = \zeta_0$)
    \item $\zetai$: Linear CCSN yield ($y_1 = \zeta_1 \log Z/\Zo$) \footnotemark{}
    \item $\zetaii$: Quadratic CCSN yield ($y_2 = \zeta_2 (\log Z/\Zo)^2$)
\end{enumerate}
The total C abundances for a linear of these yields is given from Eq.~\ref{eq:lin_combo}.
Since we are taking means of each bin, there is a statistical uncertainty associated with the mean in a given bin, $\sigma_{i, \rm model}$. The total variance of the combined yield model is 
\begin{equation}
    \sigma^2_{\rm model} = \sum_{i} \alpha_i^2 \sigma^2_{i, \rm model}
\end{equation}
Similarly, we represent the mean and variance of the observations in each bin $b$ with $\bar Z_{b, \rm obs}$ and $\sigma^2_{b, \rm obs}$.  
\begin{equation} \label{eq:mcmc_likelihood}
    \log {\cal L} \propto \chi^2 = \sum_b \frac{(\bar Z_{b,\rm\ model} - \bar Z_{b,\rm\ obs})^2} {\sigma_{b, \rm tot}^2}
\end{equation}
where the total variance in a bin is 
\begin{equation}
    \sigma_{b, \rm tot} = \sigma_{\rm int}^2 + \sigma_{b, \rm obs}^2 + \sigma^2_{b, \rm model}
\end{equation}
where we model $\sigma_{\rm int}^2$ as a free parameter in the model.
\footnotetext{In detail, $y_1$ and $y_2$ are held constant past $\log Z/\Zo < -2$ and the $y_1$ component is modelled with a positive constant offset later subtracted to avoid negative abundances.}

We adopt the following priors
\begin{subequations}
\begin{align}
    \alpha &\sim N(1, 1) \\
    \zeta^{(0)} &\sim N(2, 1) \\
    \zeta^{(1)} &\sim N(0, 1) \\
    \zeta^{(2)} &\sim N(0, 4) \\
    \ln \sigma _{\rm int}&\sim N(-3, 1)
\end{align}
\end{subequations}
We run the MCMC using Turing.jl with a No-U-turn sampler (acceptance rate of 0.65), combining 16 independent chains with 3,000 steps each for a total of 48,000 samples.

@@

\section{Results}\label{sec:results}

@@

\begin{figure*}
\centering
\includegraphics{figures/all_the_tracks.pdf}
\caption[]{
    Time evolution of gas-phase C abundances in our fiducial model for [C/Mg] versus [Mg/H] (left) and [Mg/Fe] (right).
    Each line represents a zone at a different galactic radius and is colour-coded by time. While \caah\ settles into the final trend $\sim{8}$\,Gyr ago, \caafe\ continues to evolve for much longer.
    \dbnote{radii between 2kpc and 15.5 kpc ?}
    \dbnote{add equilibrium lines (green?)}
}
\label{fig:c_evo}
\end{figure*}

@@

Fig.~\ref{fig:c_evo} shows evolutionary tracks for every zone of the fiducial model.
\caah\ trend quickly reaches our equilibrium state in \about{5}\,Gyr. \caafe\ tracks continue to evolve due to the long tail in the DTD of \ia.

The evolution of abundances in our fiducial model is as follows:
\begin{enumerate}
    \item[(1)] {\it CCSN phase.} Star formation begins. Initially, \cc\ dominate enrichment. [C/Mg] evolves with yields set by $\Ycc/\Yoc$, resulting in increasing values with [Mg/H].  [Mg/H] quickly approaches its equilibrium value, but Fe remains low in comparison (in agreement with \citealt{WAF17}).
    \item[(2)]  {\it AGB/SN Ia phase.} A few hundred million years later, delayed sources enrich the ISM. \agb\ stars release C, raising [C/Mg], and \ia\ expel Fe, lowering Mg/Fe. 
    \item[(3)] {\it Approach to equilibrium.} Several billion years later, [C/Mg] plateaus as C also approaches equilibrium. [Mg/H] reaches equilibrium. [Fe/H] continues to increase due to ongoing \ia\ from old stellar populations. 
    \item {\it Present day.} The abundances of C, Mg, and Fe reach their equilibrium values.

@@

 Fig.~\ref{fig:c_evo} {\color{red} will} show the  equilibrium tracks in \caah\, and \caafe. Based on the \citepos{WAF17} analytic models of chemical equilibrium for constant star formation, the equilibrium abundance ratio between two elements depends only on the relative yields. For C/Mg, 
\begin{equation}\label{eq:equilibrium_yields}
    {\rm [C/Mg]_{eq}} \equiv \log \left(\frac{\Ycc + \Ycagb }{y_{\rm Mg}}\right) - \log \left(\frac{Z_{\rm C,\,\odot}}{Z_{\rm Mg,\,\odot}}\right)
\end{equation}
As such, the equilibrium abundance represents the steady-state expectation for C abundances.  
However, in the \caafe\ plane, the equilibrium track is constant in [Mg/Fe] and only appears as a vertical line representing the range of equilibrium [C/Mg] values. The trends in \caafe\ are instead representative of the temporal evolution of [C/Mg].  \caafe\ is well approximated by either a single-stellar-population or a single zone model because of the shallow metallicity gradient of the Milky Way. \dbnote{is this right?}

@@

The top of Fig.~\ref{fig:zeta_f} shows models with varying strengths of the $\Ycc$ metallicity dependence, $\zcc$. With higher $\zcc$, the model predicts a steeper \caah\ trend, owing to the direct relationship between the C/Mg equilibrium abundances and yield ratios. 
However, \caafe~is minimally affected by changes to $\zcc$ for stars at a given [Mg/H].
CCSN occurs on much shorter timescales than AGB and SN Ia enrichment, so CC carbon reaches equilibrium before delayed C and Fe production modifies trends. 
\footnote{The only effects on \caafe, when considering the narrow metallicity slice, are because of either the systematic change in equilibrium abundances, the imperfect evolution of the galaxy, or that the ISM abundances are set by stars which were born at poorer metallicities. }

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Basic setup. Most papers should leave these options alone.
\documentclass[fleqn,
referee, % this line changes to double-spaced 1 column
usenatbib]{mnras}

@@

\newcommand{\fruity}{\texttt{\hyperlink{fruity}{Fruity}}}
\newcommand{\nugrid}{\texttt{\hyperlink{nugrid}{NuGrid}}}
\newcommand{\monash}{\texttt{\hyperlink{monash}{Monash}}}
\newcommand{\aton}{\texttt{\hyperlink{aton}{ATON}}}
\newcommand{\cfactor}{2.4}
\newcommand{\zetaobest}{1.90 \pm 0.10}
\newcommand{\zetaibest}{2.28 \pm 0.16}
\newcommand{\zetaiibest}{2.59 \pm 0.30}

\newcommand{\nsubgiants}{14,066}

@@

Predictions of C yields from stellar evolution and supernova (SN) models are riddled with uncertainties (see, e.g., the reviews by \citealt{romano+10, KL14}).
Some salient unknowns include mass loss rates \citep{sukhbold+16, beasor+2020}, rotational mixing \citep{frischknecht+16, LC18}, nuclear reaction rates \citep{herwig+05, herwig+austin2004}, and convection \citep{chieffi2001, ventura+13}.
Recent work additionally suggests that small changes in a stars's initial mass of 
$\sim$$0.01 \Mo$  have significant effects on the evolution and explosion outcomes of massive stars \citep{bruenn+2023, vartanyan_burrows2023}.
 Furthermore, many stars are found in binary systems \citep{sana+12}, the effects of which are almost entirely
unexplored in stellar nucleosynthesis models \citet{farmer+21}.
Accurate predictions of metal yields from stellar evolution models remain elusive.

@@

 % Folowing many previous works  (e.g. \citealt{DTS78, BF06, prantzos+18, berg+19}; see also the recent reviews by \citealt{romano22} and \citealt{RM21}.) this paper focuses on the evolution of C in GCE models.
\citet{james+24} recently argued that the enrichment history of the Galactic disk can be described as an equilibrium phenomenon.
In equilibrium GCE, trends in abundance ratios closely trace yield ratios (see their section 5.2.2).
\citet{james+23} had previously shown that the correlation between gas-phase N and O abundances (see their fig. 1) could be described as a consequence of the metallicity dependence of the relative N and O yields.
Under their equilibrium scenario, this argument should extend to any element whose enrichment by stellar populations does not require long timescales ($\gtrsim$few Gyr).
We show in section~\ref{sec:agb} that the characteristic delay times of C production should be $\sim$ 1 Gyr, within this limit. 
\dbnote{This is complicated by the difference between high alpha and low alpha, and 1-2Msun stars do produce carbon.... Also, I would like to call out reviews and general works on C at some point.}

@@

% \footnotetext{\strike{By metallicity, we mean the (mass) fraction of any element which is not H or He, denoted by $Z$. For the sun, we take $\Zo=0.0176$.} }

@@

Fig.~\ref{fig:subgiants} shows out subgiant sample plotted in the \caah\ and \caafe{} planes.\footnotemark{} In this sample, [C/Mg] increases with [Mg/H], and [C/Mg] decreases with [Mg/Fe] at fixed [Mg/H]. Our aim is to reproduce these trends.
Low-$\alpha$ stars are defined to have 
\begin{equation}\label{eq:high_alpha}
\begin{cases}
\text{[Mg/Fe]} <0.16-0.13\,\text{[Fe/H]}, & \text{[Fe/H]}<0\\
\text{[Mg/Fe]} <0.16, & \text{[Fe/H]}>0. \\
\end{cases}
\end{equation}

@@

	\begin{tabular}{l l l l l}
		\hline
         & $Z_{X,\,\sun}$ & $y_X^{\rm cc}$ & $\y_X^{\rm agb}$ & $y_X^{\rm ia}$  \\
		\hline
        C & $33.9\times10^{-4}$ & Eq.~\ref{eq:y_cc} & $\cfactor\times$\fruity &  0 \\
        O & $73.3\times10^{-4}$ & $71.3\times10^{-4}$ & 0 & 0 \\
        Mg & $6.71\times10^{-4}$ & $6.52\times 10^{-4}$ & 0 & 0 \\
        Fe & $13.7\times10^{-4}$ & $4.73\times10^{-4}$ & 0 & $7.82\times10^{-4}$ \\
        N & $10.4\times10^{-4}$ & $5\times10^{-4}$ & $5\times10^{-4}M\left(\frac{Z}{\Zo}\right)$ & 0\\
		\hline
	\end{tabular}
\end{table}

@@

\begin{table*}
	\centering
    \caption[]{For each \agb\ yield set, the IMF-averaged \agb\ C yield at solar metallicity $y_{\rm C, 0}^{\rm AGB}$, the linear fit solar yield $\zetao$ and slope $\zetai$ (see Eq.~\ref{eq:agb_z_approx}), the fraction of solar C produced in the model $f_\odot^{\rm AGB}$, and the masses and metallicities each yield table is sampled on.
    $f_{\sun}^{\rm AGB}$ is calculated based on $\zetao$ assuming a total C yield of $\Yct = 0.00275$.
    }

	\label{tab:agb}
    \begin{tabular}{c  ccc  c p{4cm} p{4cm}} % four columns, alignment for each
		\hline 
        \agb\ table 
                & $y_{\rm C, \sun}^{\rm AGB}\times10^4$ %& y
                & $\zetao^{\rm AGB}\times 10^4$ %& y
                & $\zetai^{\rm AGB}\times 10^4$
                &  $f_{\sun}^{\rm AGB}$
                & masses ($\Mo$) & metallicities ($Z$)\\
        \hline
        \fruity 
                & 3.11
                &  $3.7\pm0.3$
                & $-3.5\pm0.3$
                & 0.13
                & 1.3, 1.5, 2, 2.5, 3, 4, 5, 6
                & 0.0001, 0.0003, 0.001, 0.002, 0.003, 0.006, 0.008, 0.01, 0.014, 0.02
                \\
        \aton 
                & -0.009
                & $1.0\pm1.8$
                & $-10. \pm 3$
                & 0.04
                & 1.5, 2, 2.5, 3, 3.5, 4, 4.5, 5, 6, 6.5, 7
                & 0.0003, 0.001, 0.002, 0.004, 0.008, 0.014, 0.04
                \\
        \monash 
                &  2.96
                & $2.3 \pm 0.5$
                & $-11.0\pm 1.0$
                & 0.08
                & 1, 1.25, 1.5, 1.75, 2.25, 2.5, 2.75, 3, 3.25, 3.5, 3.75, 4, 4.5, 5, 5.5, 6, 7 
                & 0.0028, 0.007, 0.014, 0.03
                \\
        \nugrid 
                & 9.25
                & $8.1\pm 1.9$
                & $-4.6\pm1.7$
                & 0.29
                & 1, 1.65, 2, 3, 4, 5, 6, 7
                &  0.0001, 0.001, 0.006, 0.01, 0.02
                \\
		\hline
	\end{tabular}
\end{table*}

@@

Fig.~\ref{fig:y_cc} shows the integrated yields of several massive star yield models from the literature. 
\cc{} models predict a wide range of C yields, spanning nearly a factor of ten. 
 The \cite{LC18} models, which include rotation, show that the induced mixing (e.g. \citealt{frischknecht+16}) can dramatically increase the magnitude and metallicity dependence of $\Ycc$. As we will later emperically show, \cc\ C production needs to be strongly metallicity-dependent at $Z \approx \Zo$, which is consistent with \citepos{LC18} rapidly rotating models and to a lesser extent \citet{NKT13}.
The left panel of Fig.~\ref{fig:y_cc} also shows the IMF-averaged \agb\ star yields from the \fruity\ model (in Fig.~\ref{fig:agb-ssp}). Especially at $Z\approx \Zo$, most \cc\ models dominate \agb\ C production.

@@

\subsection{Supernova Type Ia}
The population averaged SN Ia yield, $y_{X}^{\rm ia}$, is given by,
\begin{equation}
    y_{X}^{\rm Ia} = m_{X}^{\rm Ia} \frac{N_{\rm Ia}}{M_\star}
\end{equation}
where $m_{X}^{\rm Ia}$ is the average mass of $X$ produced in a single SN Ia event, and $N_{\rm Ia} / M_\star$ is the number of such events per unit mass of star formation. \dbstrike{The latter can be expressed as an integral over the SN Ia delay-time-distribution (DTD), which describes the rate at which stellar populations produce X through SN Ia in \VICE.} Following \citet{james+23}, we use a $t^{-1.1}$  delay time distribution (DTD) with a minimum delay-time of 150 Myr, based on comparisons of the SN Ia rate as a function of redshift and the cosmic SFR (e.g. \citealt{maoz+mannucci12}).

SN Ia produce negligible C. Of the models from \citet{iwamoto+99}, \citet{seitenzahl+13}, and \citet{gronow+21a, gronow+21b}, the maximum C production is less than 4\% of our required solar C yield, and most likely less than 2\% (assuming a SNIa rate of $2.2\times 10^{-3} / \Mo$ \citealt{maoz+mannucci12}).

Novae are important contributes of \C[13], but do not process enough material to impact \C[12] production. \C[13] is beyond the scope of this work.

@@

As the strength of outflows controls the resulting equilibrium $\alpha$-element abundances, we extend \JJ and create a metallicity gradient by defining
\begin{equation}\label{eq:mass_loading}
\eta(R) = \frac{y_{\alpha}^{\rm cc}}{Z_{\alpha}(R)} -1 + r %+ \tau_{\star} / \tau_{\rm sfh} 
\end{equation}
where 
\begin{equation}
    \log Z_{\alpha}(R) = \log Z_{\alpha,\ \sun} + 
    0.29 + 
    \begin{cases}
        -0.015(R-5) & R < 5 \\
        -0.09(R-5) & R \geq 5
    \end{cases}
\end{equation}
is adapted from \citet{hayden+14}.  We approximate $r=0.4$. %, and $\tau_\star/\tau_{\rm sfh} = 0.0$ .
This choice of $\eta(R)$ results in a [$\alpha$/H] gradient consistent with Milky Way observations \citep[e.g.][]{hayden+14, weinberg+19, frinchaboy+13}.
If we change our assumed $y_{\rm Mg}$, the values of $\eta$ will change similarly to maintain the correct chemical trends.

@@

In GCE, total elemental abundances are (to first order) linear with respect to combinations of yields \citep[e.g.][]{WAF17}.
\dbnote{might be useful to include a complete set of differential equations for our model somewhere to better illustrate this.}
Assume that the total metallicity evolution does not depend on the elemental yield we are considering.
Say we are given a set of yield processes of $X$, $\{y_X^i\}_i$, and the resulting abundance evolution, $\{Z_{X}^i(R, t) \}_i$. Let $y_X'$, the total yield of $X$, be a linear combination of $\{y_X^i\}$, so
\begin{equation}
    y_X' = \sum_i \alpha_i\, y_X^i \quad {\rm where\ } \alpha_i \in \mathbb{R}.
\end{equation}
Then the abundance evolution of element $X$ is
\begin{equation}\label{eq:lin_combo}
    Z_X'(R, t) = \sum_i \alpha_i Z_X^i(R, t).
\end{equation}
Therefore, we can efficiently explore combinations of yields without the need for full simulations. This allows for direct observational inference of the yield combinations which best agree with the data.
We use Monte Carole Markov Chain (MCMC) inference to generate the posterior distributions of yield coefficients. Note that many other GCE parameters are nonlinear with abundance and require full multi-zone simulations, such as star formation history, outflows, and migration prescriptions. 
Also note that metallicity-dependent yields introduce a second-order effect (metallicity depends on total yields which depend on metallicity), but for small changes to total yields, this effect should be small.

Our first step is to run a multi-zone model tracking the abundances of each process we are interested in. 
We focus on the binned \textit{mean} abundances of the low-$\alpha$ sequence in \caah, and the $-0.15 \leq {\rm [M/H]} \leq -0.05$ metallicity slice in \caafe\ (with 20 bin edges between -0.4 and 0.3 in [Mg/H] and 12 bin edges between 0.025 and 0.25 in [Mg/Fe]).
We separate the C yield  into four different components.
\begin{enumerate}
    \item $\alpha$: Prefactor for an AGB yield table
    \item $\zetao$: Constant CCSN yield  ($y_0 = \zeta_0$)
    \item $\zetai$: Linear CCSN yield ($y_1 = \zeta_1 \log Z/\Zo$) \footnotemark{}
    \item $\zetaii$: Quadratic CCSN yield ($y_2 = \zeta_2 (\log Z/\Zo)^2$)
\end{enumerate}
The total C abundances for a linear of these yields is given from Eq.~\ref{eq:lin_combo}.
Since we are taking means of each bin, there is a statistical uncertainty associated with the mean in a given bin, $\sigma_{i, \rm model}$. The total variance of the combined yield model is 
\begin{equation}
    \sigma^2_{\rm model} = \sum_{i} \alpha_i^2 \sigma^2_{i, \rm model}
\end{equation}
Similarly, we represent the mean and variance of the observations in each bin $b$ with $\bar Z_{b, \rm obs}$ and $\sigma^2_{b, \rm obs}$.  We use the a likelihood
\begin{equation} \label{eq:mcmc_likelihood}
    \log {\cal L} \propto \chi^2 = \sum_b \frac{(\bar Z_{b,\rm\ model} - \bar Z_{b,\rm\ obs})^2} {\sigma_{b, \rm tot}^2}
\end{equation}
where the total variance in a bin is 
\begin{equation}
    \sigma_{b, \rm tot}^2 = \sigma_{\rm int}^2 + \sigma_{b, \rm obs}^2 + \sigma^2_{b, \rm model}.
\end{equation}
We model $\sigma_{\rm int}^2$ as a free parameter in the model.
\footnotetext{$y_1$ and $y_2$ are held constant past $\log Z/\Zo < -1$ and the $y_1$ component is modelled with a positive constant offset later subtracted to avoid negative abundances.}

We adopt the following loose priors
\begin{subequations}
\begin{align}
    \alpha &\sim N(1, 1) \\
    \zeta^{(0)} &\sim N(2, 1) \\
    \zeta^{(1)} &\sim N(0, 1) \\
    \zeta^{(2)} &\sim N(0, 4) \\
    \ln \sigma _{\rm int}&\sim N(-3, 1)
\end{align}
\end{subequations}
We run the MCMC using Turing.jl \citep{turingjl} with a No-U-turn sampler (adaptive Hamilton Monte-Carlo algorithm, \citealt{NUTS}) with an acceptance rate of 0.65. We combine 16 independent chains with 3,000 steps each for a total of 48,000 samples.

@@

The median parameters derived from this model provide the yield choices for the fiducial and each AGB multizone model shown later.
\section{Results}\label{sec:results}

@@

\begin{figure*}
\centering
\includegraphics{figures/all_the_tracks.pdf}
\caption[]{
    Time evolution of gas-phase C abundances in our fiducial model for [C/Mg] versus [Mg/H] (left) and [Mg/Fe] (right).
    Each line represents a zone at a different galactic radius and is colour-coded by time. While \caah\ settles into the final trend $\sim{8}$\,Gyr ago, \caafe\ continues to evolve for much longer.
    Only zones between 2 and 15.5 kpc are shown. The green lines represent the equilibrium tracks.
}
\label{fig:c_evo}
\end{figure*}

@@

Fig.~\ref{fig:c_evo} shows evolutionary tracks for every zone of the fiducial model.
\caah\ trend quickly reaches an equilibrium state in \about{5}\,Gyr. \caafe\ tracks continue to evolve due to the long tail in the DTD of \ia.

The evolution of abundances in our fiducial model is as follows:
\begin{enumerate}
    \item[(1)] {\it CCSN phase.} Star formation begins. Initially, \cc\ dominate enrichment. [C/Mg] evolves with yields set by $\Ycc/\Yoc$, resulting in increasing values with [Mg/H].  [Mg/H] quickly approaches its equilibrium value, but Fe remains low in comparison (in agreement with \citealt{WAF17}).
    \item[(2)]  {\it AGB/SN Ia phase.} A few hundred million years later, delayed sources enrich the ISM. \agb\ stars release C, raising [C/Mg], and \ia\ expel Fe, lowering Mg/Fe. 
    \item[(3)] {\it Approach to equilibrium.} Several billion years later, [C/Mg] plateaus as C also approaches equilibrium. [Mg/H] reaches equilibrium. [Fe/H] continues to increase due to ongoing \ia\ from old stellar populations. 
    \item[(4)] {\it Present day.} The abundances of C, Mg, and Fe reach their equilibrium values.

@@

 Fig.~\ref{fig:c_evo} also shows the  equilibrium tracks in \caah\, and \caafe. Based on the \citepos{WAF17} analytic models of chemical equilibrium for constant star formation, the equilibrium abundance ratio between two elements depends only on the relative yields. For C/Mg, 
\begin{equation}\label{eq:equilibrium_yields}
    {\rm [C/Mg]_{eq}} \equiv \log \left(\frac{\Ycc + \Ycagb }{y_{\rm Mg}}\right) - \log \left(\frac{Z_{\rm C,\,\sun}}{Z_{\rm Mg,\,\sun}}\right)
\end{equation}
As such, the equilibrium abundance represents the steady-state expectation for C abundances.  
However, in the \caafe\ plane, equilibrium [Mg/Fe] is single-valued, so equilibrium track is a vetical line representing the range of ${\rm [C/Mg]_{eq}}$ across metallicity. The trends in \caafe\ are instead representative of the temporal evolution of [C/Mg].  \caafe\ is well approximated by either a single-stellar-population or a single zone model because of the shallow metallicity gradient of the Milky Way.

@@

The top of Fig.~\ref{fig:zeta_f} shows models with varying strengths of the $\Ycc$ metallicity dependence, $\zcc$. With higher $\zcc$, the model predicts a steeper \caah\ trend, owing to the direct relationship between the C/Mg equilibrium abundances and yield ratios. 
However, \caafe~is minimally affected by changes to $\zcc$ for stars at a given [Mg/H].
CCSN occurs on much shorter timescales than AGB and SN Ia enrichment, so CCSN carbon reaches equilibrium before delayed C and Fe production modifies trends. 
\footnote{The only effects on \caafe, when considering the narrow metallicity slice, are because of either the systematic change in equilibrium abundances, the imperfect evolution of the galaxy, or that the ISM abundances are set by stars which were born at poorer metallicities. }
